---
title: "Project 1 - Group E - JSON"
author: "Kristin Calvert"
date: "6/5/2019"
output: 
  html_document:
    toc:  TRUE
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(jsonlite)
library(tidyverse)
library(knitr)
library(tidyr)

```

# Project 1 - Group # - JSON

# **J**ava**S**cript **O**bject **N**otation

## What is JSON?

JSON is a data-interchange language. While originally derived from JavaScript, JSON is programming language neutral. It consists of human readable colletions of name/value pairs
It is "less verbose" than XML.^2^ Where XML allows for data and metadata and commenting, JSON objects are expressed as name-value pairs. Though this limits the data JSON is able to communicate, it does allow for straigtforward reading of the data by programs like R. 

## Basic Structure

The JSON data format denotes an object as a comma-separated list of properties contained within braces. JSON supports six types of values: number, string, Boolean, array, object, `NULL`. Whitespace is ignored between elements.^1^

### Sample JSON Object

Sample Object^3^ 

```{r JSONsample, echo=TRUE, eval=FALSE}
    "actor": {
      "name": "Tom Cruise",
      "age": 56,
      "Born At": "Syracus, NY",
      "Birthdate": "July 3 1962",
      "photo": "https://jsonformatter.org/img/tom-cruise.jpg"
    }
```

## Why is it good for storing data?

*Speed*
JSON is a pure data format designed for transmission. It is compact; the absence of metadata, schema, or other information make JSON faster to parse compared to XML or other languages because JSON uses less data overall. 

*Readability*
Requires additional validation as small mistakes in coding commas or parenthes cause issues. Lacks schema. Less flexible, which simplifies code.^1^ 

*Compatibility*
JSON is language neutral. Additionally, it uses conventions familiar to other programming languages. It easily integrated in web resources, especially JavaScript. Many notable public APIs use JSON.^4^ 

### APIs supporting JSON

Data format | API^5^
-------------|----------------
JSON only | Facebook Graph API, Google Maps API, Twitter API, AccuWeather API, Pinterest API, Reddit API, Foursquare API
JSON and XML | Google Cloud Storage, Linkedin API, Flickr API


References:

^1^ [JSON.org](http://www.json.org/). Accessed 2019-06-12

^2^ [Friesen J. (2016) Introducing JSON. In: Java XML and JSON. Apress, Berkeley, CA](https://link-springer-com.prox.lib.ncsu.edu/chapter/10.1007/978-1-4842-1916-4_7)

^3^ [JSON Example with Data Types Including JSON Array, on CODEBLOGMONEY.com](https://codeblogmoney.com/json-example-with-data-types-including-json-array/)

^4^ [Why you should be using JSON vs XML by Vineet Joshi. Posted Jan 18, 2017 from Cloud-Elements](https://blog.cloud-elements.com/using-json-over-xml)

^5^ [XML vs JSON by Tom Strassner](https://www.cs.tufts.edu/comp/150IDS/final_papers/tstras01.1/FinalReport/FinalReport.html)


## Packages for JSON data

There are three packages for importing JSON data into R: [rjson](https://cran.r-project.org/web/packages/rjson/index.html), [rjsonio](https://cran.r-project.org/web/packages/RJSONIO/index.html), and [jsonite](https://cran.r-project.org/web/packages/jsonlite/index.html).

The packages function similarly; however, there are differences in how the parse the data, indicate missing values, and contain options for controlling how data is presented or simplified. `rjson` has fewer options for the latter. 

We have selected `jsonlite` for this project. The data set we will use for the analysis has nested data frames. This package allows us to flatten the data into a non-nested data frame.


# CORGIS Crime Dataset

We have selected to use the County Crime dataset available from [CORGIS Dataset Project](https://think.cs.vt.edu/corgis/json/county_crime/county_crime.html). CORGIS is a project out of Virginia Tech to provide clean, simple to use datasets for people learning to code. They provide data across a number of subject areas and in a variety of common data formats. 

The County Crime dataset originates from the Unified Crime Reporting Statistics coming out of U.S. Justice Department and the Federal Bureau of Investigations. The data include crime rates by county in the United States. The data are broken down into categories which broadly fall under either property or violent crime designations. Data are reported from "city law enforcement agencies 10,000 and over in population and county law enforcement agencies 25,000 and over in population." !https://www.ucrdatatool.gov/faq.cfm Geographic data come from the U.S. Census Bureau.

The data include two character variables:

+ Department: the law enforment agency providing the data
+ State: the U.S. state where the department is located.

and 20 numeric variables. The numeric variables are structured as data frames within the larger data set.

+ Violent crime rates: in 5 categories
  - All
  - Murder
  - Rape
  - Robbery
  - Assault
+ Property crime rates: in 4 categories
  - Burglary
  - Larceny
  - All
  - Motor
+ Violent crime total offenses: in 5 categories (same as above)
+ Property crime total offenses: in 4 categories (same as above)
+ Year: the reporting year for the data

```{r readDS}
#Use the flatten option to remove the nested data frames
county_crime <- fromJSON("https://think.cs.vt.edu/corgis/json/county_crime/county_crime.json", flatten = TRUE)

```


## Data Analysis

### Contingency Table

The combined crime data includes reports from different law enformement agencies. In North Carolina, the data primarily come from the county sheriffs and municipal police departments. See table below. The "other" category includes municipalities that did not include an agency name and a department of public safety. 

```{r contingencyTables}

#Limit data to NC. Create a new variable that indicate the agency type, based on a basic string search for terms.
north_carolina_crime <- filter(county_crime, county_crime$State == 'North Carolina')
north_carolina_crime$Department_Type <-(ifelse(grepl("Police",north_carolina_crime$Department),"Police", ifelse(grepl("Sheriff", north_carolina_crime$Department),"Sheriff","Other")))

table(north_carolina_crime$Department_Type) %>% kable(col.names = c("Agency Type", "Frequency"), caption = "Table 1. Types of Reporting Agencies in North Carolina")

```

### Bar Charts

We can also start with some simple bar charts to get a sense of the data. We can examine which states have the highest numbers of report crimes (see Figure 1). Another simple analysis that can be done with the data is to look at trends over time. We have taken a snippet of data from North Carolina for 2010 - 2012 to get an overview of the change in number of property crimes broken out by type (see Figure 2). The visual inspection of the data may reveal interesting trends we would want to examine in more detail. 

```{r stateCrime}

#Subset data by year = 2010 and get the total number of crimes by state
state_totals <- county_crime %>% filter(Year == 2010) %>% group_by(State)%>% transmute(total_violent = sum(Data.Totals.Violent.All), total_property = sum(Data.Totals.Property.All), total_crime = (total_violent+total_property)) %>% distinct()
state_totals <- state_totals %>% arrange(desc(state_totals$total_crime))

#Plot the total number of crimes committed in each state
state_bar <- ggplot(state_totals, aes(x = reorder(State, -total_crime), y = total_crime))
state_bar + geom_bar(stat = "Identity", color = "purple") + xlab(element_blank()) + ylab("Total Crimes Reported") + ggtitle("Figure 1. Number of Crimes Reported in Each State, 2010")+ scale_y_continuous(labels = scales::comma) + theme(axis.text.x = element_text(size = 6, angle = 45))

north_carolina_property_crime <- north_carolina_crime %>% select("Department", "Data.Totals.Property.Burglary", "Data.Totals.Property.Larceny", "Data.Totals.Property.Motor","Year")
north_carolina_property_crime <- north_carolina_property_crime %>% gather(key = type, value = totals, 2:4)

bar_plot <- ggplot(subset(north_carolina_property_crime, Year >= 2010 & Year <= 2015), aes(x = type, y = totals, fill = as.factor(Year)))
bar_plot + stat_summary(fun.y = "sum", geom = "bar", position = "dodge") + scale_x_discrete(labels = c("Burglary","Larceny","Motor")) + xlab("Property Crime Category") + ylab("Totals") +scale_fill_discrete(element_blank()) + ggtitle("Figure 2. Property Crime Rates in North Carolina, 2010-2012")

```

### Scatter Plots

One area of interest could be examining the differences in crime rates based on the population of the county. We could do analyses to determine if there is a statistically significant correlation between county size and crime rates. We have created a new variable that classifies each county into a size factor. Then we plotted the summary data for each level on the box plot. There are a huge number of outliers that could be worth investigating further. Our population sizes were chosen arbitrarily; a proper analysis would need to include a justification for each group. Even so we see a slight increase in the mean crime rate in larger population centers. 

```{r populationSize}
#Create a new factor that assigns a size category to each report based on the population of the county. Population figures chosen at arbitrary intervals.

county_crime$county_size <- ifelse(county_crime$Data.Population > 500000, "Very Large", ifelse(county_crime$Data.Population >= 100000, "Large", ifelse(county_crime$Data.Population >= 50000, "Mid-size", ifelse(county_crime$Data.Population >15000, "Small", ifelse(county_crime$Data.Population >0,"Tiny","Missing")))))
county_crime$county_size <- factor(county_crime$county_size, levels = c("Missing","Tiny","Small","Mid-size","Large","Very Large"))

#Create boxplots for each size. Counties with zero population were excluded.
box_plot <- ggplot(subset(county_crime, Year == 2010 & county_size != "Missing"))
box_plot + geom_boxplot(aes(y = Data.Rates.Property.All, fill = county_size)) + ylab("Property Crime Rates") + theme(axis.text.x = element_blank()) + scale_fill_discrete(name = "County Size") + ggtitle("Figure 3. 2010 US Property Crime Rates by County Size")

```

### Scatter

Again, with our interest in population size and crime rate, we have plotted the overall reported rates of violent crime for 2010 for populations greater than zero and less than 2 million. We don't see an obvious relationship, but we could iterate a linear model to the data across years to determine if there could be a relationship.

```{r scatter}
scatter_plot <- ggplot(subset(county_crime, Year == 2010 & Data.Population > 0 & Data.Population < 2000000))
scatter_plot + geom_point(aes(x = Data.Population, y = Data.Rates.Violent.All), colour = "red") + ggtitle("2010 US Violent Crime Rates by Population (under 2M)") + ylab("Total Violent Crime Rates") + xlab("Population") + scale_x_continuous(labels = scales::comma)

```

